---
title: "exercies of R- Tidyverse"
author: "Evrim Bilgen"
date: "3/5/2021"
output: html_document
references: DataCamp
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction to the Tidyverse
## Data wrangling

 filter for particular observations, arrange the observations in a desired order, and mutate to add or change a column
```{r gapminder}
library(gapminder)
library(dplyr)

# Use mutate to change lifeExp to be in months

gapminder %>%
  mutate(lifeExp = 12 * lifeExp)


# Use mutate to create a new column called lifeExpMonths

gapminder %>% 
  mutate(lifeExpMonths= 12 * lifeExp)
```

### Combining filter, mutate, and arrange

I am gonna combine all three of the verbs you've learned in this chapter, to find the countries with the highest life expectancy, in months, in the year 2007.

```{r tidyverse}
library(gapminder)
library(dplyr)

# Filter, mutate, and arrange the gapminder dataset


gapminder %>% mutate(lifeExpMonths= 12*lifeExp) %>% 
   filter(year==2007) %>% 
     arrange(desc(lifeExpMonths))

```

##Data visualization

###Comparing population and GDP per capita

I am changing the scatter plot of gapminder_1952 so that (pop) is on the x-axis and GDP per capita (gdpPercap) is on the y-axis

```{r visu}
library(gapminder)
library(dplyr)
library(ggplot2)

gapminder_1952 <- gapminder %>%
  filter(year == 1952)

# Change to put pop on the x-axis and gdpPercap on the y-axis
ggplot(gapminder_1952, aes(x = pop, y = gdpPercap)) +
  geom_point()
```
  
###Putting the x- and y- axes on a log scale

Creating a scatter plot with population (pop) on the x-axis and GDP per capita (gdpPercap) on the y-axis. Put both the x- and y- axes on a log scale

```{r visu2}
library(gapminder)
library(dplyr)
library(ggplot2)

gapminder_1952 <- gapminder %>%
  filter(year == 1952)

# Scatter plot comparing pop and gdpPercap, with both axes on a log scale

ggplot(gapminder_1952, aes(x = pop, y = gdpPercap)) +
  geom_point() + scale_x_log10() + scale_y_log10()
```


###Adding size and color to a plot

Modifying  the scatter plot so that the size of the points represents each country's GDP per capita (gdpPercap)


```{r visu3}

library(gapminder)
library(dplyr)
library(ggplot2)

gapminder_1952 <- gapminder %>%
  filter(year == 1952)

# Add the size aesthetic to represent a country's gdpPercap
ggplot(gapminder_1952, aes(x = pop, y = lifeExp, color = continent, size = gdpPercap)) +
  geom_point() +
  scale_x_log10()

```

###Creating a subgraph for each continent

Create a scatter plot of gapminder_1952 with the x-axis representing population (pop), the y-axis representing life expectancy (lifeExp), and faceted to have one subplot per continent (continent). Put the x-axis on a log scale.

```{r visu4}
library(gapminder)
library(dplyr)
library(ggplot2)

gapminder_1952 <- gapminder %>%
  filter(year == 1952)

# Scatter plot comparing pop and lifeExp, faceted by continent

ggplot(gapminder_1952, aes(x=pop, y=lifeExp))+
geom_point()+
scale_x_log10()+
facet_wrap(~ continent)

```

##Grouping and summarizing

###Summarizing by continent

Filter the gapminder data for the year 1957. Then find the median life expectancy (lifeExp) and maximum GDP per capita (gdpPercap) within each continent, saving them into medianLifeExp and maxGdpPercap, respectively


```{r grouping1}
library(gapminder)
library(dplyr)

# Find median life expectancy and maximum GDP per capita in each continent in 1957

 gapminder %>% filter(year==1957) %>%
 group_by(continent) %>%
   summarize(medianLifeExp=median(lifeExp), maxGdpPercap=max(gdpPercap) ) 


```

###Visualizing median GDP per capita per continent over time

* Summarize the gapminder dataset by continent and year, finding the median GDP per capita (gdpPercap) within each and putting it into a column called medianGdpPercap. Use the assignment operator <- to save this summarized data as by_year_continent

* Create a scatter plot showing the change in medianGdpPercap by continent over time. Use color to distinguish between continents, and be sure to add expand_limits(y = 0) so that the y-axis starts at zero.

```{r grouping2}
library(gapminder)
library(dplyr)
library(ggplot2)

# Summarize medianGdpPercap within each continent within each year: by_year_continent



by_year_continent <- gapminder %>%
  group_by(continent,year) %>%
  summarize(medianGdpPercap = median(gdpPercap))

# Create a scatter plot showing the change in medianLifeExp over time

ggplot(by_year_continent, aes(x = year, y = medianGdpPercap, color=continent)) + 
 geom_point() + 
 expand_limits(y = 0)


```


##Types of visualizations

###Line Plots

* Use group_by() and summarize() to find the median GDP per capita within each year and continent, calling the output column medianGdpPercap. Use the assignment operator <- to save it to a dataset called by_year_continent.

* Use the by_year_continent dataset to create a line plot showing the change in median GDP per capita over time, with color representing continent. Be sure to use expand_limits(y = 0) to include 0 on the y-axis.

```{r lp}
library(gapminder)
library(dplyr)
library(ggplot2)

# Summarize the median gdpPercap by year & continent, save as by_year_continent

by_year_continent <- gapminder %>%
  group_by(year,continent) %>%
  summarize(medianGdpPercap = median(gdpPercap))

  
# Create a line plot showing the change in medianGdpPercap by continent over time

ggplot(by_year_continent, aes(x = year, y = medianGdpPercap, color=continent)) + geom_line() + 
  expand_limits(y = 0)

# Create a line plot showing the change in medianGdpPercap over time
ggplot(by_year_continent, aes(x = year, y = medianGdpPercap)) + geom_line() + 
  expand_limits(y = 0)
```


###Bar Plots

* Filter for observations in the Oceania continent in the year 1952. Save this as oceania_1952.

* Use the oceania_1952 dataset to create a bar plot, with country on the x-axis and gdpPercap on the y-axis.

```{r bp}
library(gapminder)
library(dplyr)
library(ggplot2)

# Filter for observations in the Oceania continent in 1952

# Filter for observations in the Oceania continent in 1952
oceania_1952 <- gapminder %>%
  filter(year == 1952, continent == "Oceania")

# Create a bar plot of gdpPercap by country
ggplot(oceania_1952, aes(x = country, y = gdpPercap)) + geom_col()

```



###Histograms

* Use the gapminder_1952 dataset (code is provided) to create a histogram of country population (pop) in the year 1952, putting the x-axis on a log scale with scale_x_log10().


```{r his}
library(gapminder)
library(dplyr)
library(ggplot2)

gapminder_1952 <- gapminder %>%
  filter(year == 1952)

# Create a histogram of population (pop), with x on a log scale

ggplot(gapminder_1952, aes(x = pop)) + 
  geom_histogram() + 
  scale_x_log10()

```


###Boxplots

* Use the gapminder_1952 dataset (code is provided) to create a boxplot comparing GDP per capita (gdpPercap) among continents. Put the y-axis on a log scale with scale_y_log10()

* Add a title to the graph: Comparing GDP per capita across continents

```{r bp}

library(gapminder)
library(dplyr)
library(ggplot2)

gapminder_1952 <- gapminder %>%
  filter(year == 1952)

# Add a title to this graph: "Comparing GDP per capita across continents"
ggplot(gapminder_1952, aes(x = continent, y = gdpPercap)) +
  geom_boxplot() +
  scale_y_log10() +
  ggtitle("Comparing GDP per capita across continents")

```



